---
title: "Combined report on missings patterns."
author: "Piotr Fic"
date: "11 08 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir=normalizePath('../../../'))
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r packages}
library(ggplot2)
library(ggthemes)
library(dplyr)
library(scales)
```

```{r data_reading}
data <- read.csv("datasets_store/patterns_analysis/readed_json_info.csv")[, -1]
selected_df <- read.csv("datasets_store/datasets_selection/selected_datasets.csv")
data <- data[data$id %in% selected_df$ID, ]
data <- data[!is.na(data$name), ]
patterns <- read.csv(file = "datasets_store/patterns_analysis/summary_of_patterns_in_all.csv")
patterns <- patterns[patterns$id %in% data$id, ]
```

### Data used.

- The report was created on the basis of information on 132 datasets, both from the regression and classification tasks.

## Key information about distribution of missings.

### *Figure 1*

```{r percentage_of_missings}
to_plot <- data
to_plot$per_missings <- round(100*to_plot$no_missings/(to_plot$no_instances*to_plot$no_features), digits = 1)

ggplot(to_plot, aes(per_missings))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of missings")+
  ylab("% of datasets")+
  ggtitle("Histogram of percentage of missings among datasets.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.7, 0.1), labels = seq(0, 70, 10))+
  theme_linedraw()
```

- Most of the datasets have missings percentage lower than 5%.
- About 25% of datasets have a fraction of missings between 5% to 20%.
- Higher percentages of missings occur in a small number of datasets.

### *Figure 2*

```{r percentage_instances_missings}
to_plot <- data
to_plot$per_instances_miss <- round(100*to_plot$no_instances_w_miss/to_plot$no_instances)

ggplot(to_plot, aes(per_instances_miss))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of instances with missings")+
  ylab("% of datasets")+
  ggtitle("Histogram of percentage of instances with missing values.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.3, 0.1), labels = seq(0, 30, 10), limits = c(0, 0.3))+
  theme_linedraw()
```

- There are two dominant groups of occurrence of missings among instances.
- The first group contains datasets where the percentage of instances with missings is lower than 15%.
- The second group is formed by datasets with a very high percentage of instances with missings (above 90%).

## Key information about distribution of missings patterns.

### *Figure 3*

```{r patterns_in_instances_fraction}
to_plot <- data
to_plot$fraction <- round(100*to_plot$no_of_patterns/to_plot$no_instances, digits = 1)

ggplot(to_plot, aes(fraction))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("no_of_patterns/no_of_instances [%]")+
  ylab("% of datasets")+
  ggtitle("Histogram of fraction of unique patterns to number of instances.")+
  scale_x_continuous(breaks = seq(0, 100, by = 5))+
  scale_y_continuous(breaks = seq(0, 0.8, 0.1), labels = seq(0, 80, 10), limits = c(0, 0.75))+
  theme_linedraw()
```

- The figure above presents a dispersion of missings' pattern among datasets.
- Most of the datasets have patterns which have less than 10% of unique combinations compared to  
the number of instances in datasets.
- There are very few datasets which have diverse patterns with a unique combination for most rows of the dataset.

### *Figure 4*

```{r variables_in_patterns}
ggplot(patterns, aes(var_per))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of variables in pattern")+
  ylab("% of patterns")+
  ggtitle("Histogram of percentage of variables in patterns.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.60, 0.10), labels = seq(0, 60, 10))+
  theme_linedraw()
```

- Percentage of variables in the pattern means a fraction of variables in dataset's row which have missing values.  
Different combinations of variables with missings create missings patterns.
- Most of the missings patterns contain less than 10% of variables.
- The second noticeable, but small group create patterns with the percentage of variables between 65% to 75%.

### *Figure 5*

```{r average_vars_in_patterns_by_dataset}
group_by(patterns, id)%>%
  summarise("average" = mean(var_per))%>%
ggplot(aes(average))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("average % of variables in pattern")+
  ylab("% of datasets")+
  ggtitle("Histogram of average percentage of variables in pattern by datasets.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.3, 0.1), labels = seq(0, 30, 10), limits = c(0, 0.3))+
  theme_linedraw()

```

- On the figure above the percentage of variables in patterns is averaged for every dataset.
- Averaging by datasets introduce more diversity in distribution.
- Conclusions remain similar, most of the datasets have patterns which on average contain less than 25% of variables.

### *Figure 6*

```{r percentage_instances_in_patterns}
ggplot(patterns, aes(x = row_per))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, max(patterns$row_per), 2))+
  xlab("% of instances in pattern")+
  ylab("% of patterns")+
  ggtitle("Histogram of percentage of instances in patterns.")+
  scale_x_continuous(breaks = seq(0, max(patterns$row_per), 4))+
  scale_y_continuous(breaks = seq(0, 1, 0.25), labels = seq(0, 100, 25))+
  theme_linedraw()
```

- Almost all patterns seem to cover an only very small fraction of dataset's instances.

### *Figure7*

```{r average_percentage_instances_pattern}
group_by(patterns, id)%>%
  summarise("average" = mean(row_per))%>%
ggplot(aes(x = average))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, max(patterns$row_per), 2))+
  xlab("average % of instances in pattern")+
  ylab("% of datasets")+
  ggtitle("Histogram of average percentage of instances in pattern by datasets.")+
  scale_x_continuous(breaks = seq(0, max(patterns$row_per), 4))+
  scale_y_continuous(breaks = seq(0, 0.5, 0.1), labels = seq(0, 50, 10))+
  theme_linedraw()
```

- Calculating the average of a number of instances in patterns for each dataset help to see more diversity.
- Still for most datasets a single pattern on average covers below 4% of instances.
- However, some datasets have missings patters which on the average cover more than half of the instances in the dataset.

<br></br>
<br></br>