---
title: "Combined report on missings patterns."
author: "Piotr Fic"
date: "11 08 2020"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: 150px;
  margin-right: 150px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir=normalizePath('../../../'))
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r packages}
library(ggplot2)
library(ggthemes)
library(dplyr)
library(scales)
library(patchwork)
library(kableExtra)
```

```{r data_reading}
data <- read.csv("datasets_store/patterns_analysis/readed_json_info.csv")[, -1]
selected_df <- read.csv("datasets_store/datasets_selection/selected_datasets.csv")
data <- data[data$id %in% selected_df$ID, ]
data <- data[!is.na(data$name), ]
patterns <- read.csv(file = "datasets_store/patterns_analysis/summary_of_patterns_in_all.csv")
patterns <- patterns[patterns$id %in% data$id, ]
```

### Data used.

- The report was created on the basis of information on 132 datasets, both from the regression and classification tasks.

## Key information about distribution of missings.

### *Figure 1*

```{r percentage_of_missings}
data$per_missings <- round(100*data$no_missings/(data$no_instances*data$no_features), digits = 1)

p1 <- ggplot(data, aes(per_missings))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of missings")+
  ylab("% of datasets")+
  ggtitle("Histogram of percentage of missings among datasets.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.7, 0.1), labels = seq(0, 70, 10))+
  theme_linedraw()
p1
```

- Most of the datasets have missings percentage lower than 5%.
- About 25% of datasets have a fraction of missings between 5% to 20%.
- Higher percentages of missings occur in a small number of datasets.

### *Figure 2*

```{r percentage_instances_missings}
data$per_instances_miss <- round(100*data$no_instances_w_miss/data$no_instances)

p2 <- ggplot(data, aes(per_instances_miss))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of instances with missings")+
  ylab("% of datasets")+
  ggtitle("Histogram of percentage of instances with missing values.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.3, 0.1), labels = seq(0, 30, 10), limits = c(0, 0.3))+
  theme_linedraw()
p2
```

- There are two dominant groups of occurrence of missings among instances.
- The first group contains datasets where the percentage of instances with missings is lower than 15%.
- The second group is formed by datasets with a very high percentage of instances with missings (above 90%).

### *Figure 3*

```{r percentage_variables_missings}
data$per_variables_miss <- round(100*(data$no_cat_w_miss+data$no_num_w_miss)/data$no_features)

p3 <- ggplot(data, aes(per_variables_miss))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of variables with missings")+
  ylab("% of datasets")+
  ggtitle("Histogram of percentage of variables with missing values.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.2, 0.1), labels = seq(0, 20, 10), limits = c(0, 0.2))+
  theme_linedraw()
p3
```

## Key information about distribution of missings patterns.

### *Figure 4*

```{r patterns_in_instances_fraction}
data$fraction <- round(100*data$no_of_patterns/data$no_instances, digits = 1)

ggplot(data, aes(fraction))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("no_of_patterns/no_of_instances [%]")+
  ylab("% of datasets")+
  ggtitle("Histogram of fraction of unique patterns to number of instances.")+
  scale_x_continuous(breaks = seq(0, 100, by = 5))+
  scale_y_continuous(breaks = seq(0, 0.8, 0.1), labels = seq(0, 80, 10), limits = c(0, 0.75))+
  theme_linedraw()
```

- The figure above presents a dispersion of missings' pattern among datasets.
- Most of the datasets have patterns which have less than 10% of unique combinations compared to  
the number of instances in datasets.
- There are very few datasets which have diverse patterns with a unique combination for most rows of the dataset.

### *Figure 5*

```{r variables_in_patterns}
ggplot(patterns, aes(var_per))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("% of variables in pattern")+
  ylab("% of patterns")+
  ggtitle("Histogram of percentage of variables in patterns.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.60, 0.10), labels = seq(0, 60, 10))+
  theme_linedraw()
```

- Percentage of variables in the pattern means a fraction of variables in dataset's row which have missing values.  
Different combinations of variables with missings create missings patterns.
- Most of the missings patterns contain less than 10% of variables.
- The second noticeable, but small group create patterns with the percentage of variables between 65% to 75%.

### *Figure 6*

```{r average_vars_in_patterns_by_dataset}
p6 <- group_by(patterns, id)%>%
  summarise("average" = mean(var_per))%>%
ggplot(aes(average))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, 100, 5))+
  xlab("average % of variables in pattern")+
  ylab("% of datasets")+
  ggtitle("Histogram of average percentage of variables in pattern by datasets.")+
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.3, 0.1), labels = seq(0, 30, 10), limits = c(0, 0.3))+
  theme_linedraw()
p6
```

- On the figure above the percentage of variables in patterns is averaged for every dataset.
- Averaging by datasets introduce more diversity in distribution.
- Conclusions remain similar, most of the datasets have patterns which on average contain less than 25% of variables.

### *Figure 7*

```{r percentage_instances_in_patterns}
ggplot(patterns, aes(x = row_per))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, max(patterns$row_per), 2))+
  xlab("% of instances in pattern")+
  ylab("% of patterns")+
  ggtitle("Histogram of percentage of instances in patterns.")+
  scale_x_continuous(breaks = seq(0, max(patterns$row_per), 4))+
  scale_y_continuous(breaks = seq(0, 1, 0.25), labels = seq(0, 100, 25))+
  theme_linedraw()
```

- Almost all patterns seem to cover an only very small fraction of dataset's instances.

### *Figure 8*

```{r average_percentage_instances_pattern}
group_by(patterns, id)%>%
  summarise("average" = mean(row_per))%>%
ggplot(aes(x = average))+
  geom_histogram(aes(y = stat(count/sum(count))), fill='#2b8cbe', breaks = seq(0, max(patterns$row_per), 2))+
  xlab("average % of instances in pattern")+
  ylab("% of datasets")+
  ggtitle("Histogram of average percentage of instances in pattern by datasets.")+
  scale_x_continuous(breaks = seq(0, max(patterns$row_per), 4))+
  scale_y_continuous(breaks = seq(0, 0.5, 0.1), labels = seq(0, 50, 10))+
  theme_linedraw()
```

- Calculating the average of a number of instances in patterns for each dataset help to see more diversity.
- Still for most datasets a single pattern on average covers below 4% of instances.
- However, some datasets have missings patters which on the average cover more than half of the instances in the dataset.

<br></br>
<br></br>

## Decision tree

Based on the above analysis, we can try to group data sets against key statistics.

```{r tree_setup, include=FALSE}
source("datasets_store/patterns_analysis/decision_tree.R")

#Modification of figures to use with tree plots
p1 <- p1+ggtitle("Percentage of missings")
p2 <- p2+ggtitle("Instances with missings")
p3 <- p3+ggtitle("Variables with missings")
p6 <- p6+ggtitle("Variables in patterns")

#Decision variables to choose:
# - "per_missings", "per_variables_missings", "per_instances_missings", "patterns_fraction", "avg_per_var_pattern", "avg_per_inst_pattern"

leaves_stats <- function(df, cols){
  result <- list()
  for (i in (1:max(df$group_id))) {
    
  avg <- filter(df, group_id == i)%>%
    summarise_all(mean)
  
  st_dev <- filter(df, group_id == i)%>%
    summarise_all(sd)
  
  avg <- t(avg[, -c(1,2,3)])
  st_dev <- t(st_dev[, -c(1,2,3)])
  #row.names(avg) <- c("%_missings", "%_vars_missings", "%_inst_missings", "%_patterns", "%_vars_patterns", "%_inst_patterns")
  #row.names(st_dev) <- c("%_missings", "%_vars_missings", "%_inst_missings", "%_patterns", "%_vars_patterns", "%_inst_patterns")
  row.names(avg) <- colnames(df[-c(1,2,3)])
  row.names(st_dev) <- colnames(df[-c(1,2,3)])
  
  temp <- data.frame(avg, st_dev)
  temp <- round(temp, 1)
  colnames(temp) <- c(paste(i, "avg", sep = "_"), paste(i, "sd ", sep = "_"))
  colnames(temp) <- cell_spec(colnames(temp), background = cols[i])
  
  if(i>1){
    row.names(temp) <- NULL
    result[[i]] <- temp
  }else{
    row.names(temp) <- colnames(df[-c(1,2,3)])
    result[[i]] <- temp
  }
  }
  return(result)
}
```

### *Tree 1*

Splitting variables:

- percentage of missings [10]
- percentage of instances with missings [15, 90]
- fraction of patterns in instances [10]

<center>
```{r tree1, fig.width=12}
decisions1 <- list(
  "per_missings" = 10,
  "per_instances_missings" = c(15, 90),
  "patterns_fraction" = 10
)

tree1 <- decision_tree(df, decisions1)
plot(tree1$tree)
```
</center>

### *Tree 2*

Splitting variables:

- percentage of missings [5]
- percentage of instances with missings [35]
- fraction of patterns in instances [10]

<center>
```{r tree2, fig.width=12}
decisions2 <- list(
  "per_missings" = 5,
  "per_instances_missings" = 35,
  "patterns_fraction" = 10
)

tree2 <- decision_tree(df, decisions2)
plot(tree2$tree)
```
</center>

### *Tree 3*

Splitting variables:

- percentage of missings [5]
- percentage of instances with missings [35]
- fraction of patterns in instances [10]

<center>
```{r tree3, fig.width=12}
decisions3 <- list(
  "per_missings" = 5,
  "per_instances_missings" = 35,
  "patterns_fraction" = 10
)

tree3 <- decision_tree(df, decisions3)
plot(tree3$tree)
```
</center>

### *Tree 4*

Splitting variables:

- percentage of missings [5]
- percentage of instances with missings [15]
- average percentage of variables in dataset's patterns [10, 25]

<center>
```{r tree4, fig.width=12, fig.height=3}
decisions4 <- list(
  "per_missings" = 5,
  "per_instances_missings" = 15,
  "avg_per_var_pattern" = c(10, 25)
)

p1 <- p1+geom_vline(xintercept = decisions4$per_missings, color = "#fc3d03", linetype = "twodash", size = 1.25)
p2 <- p2+geom_vline(xintercept = decisions4$per_instances_missings, color = "#fc3d03", linetype = "twodash", size = 1.25)
p6 <- p6+geom_vline(xintercept = decisions4$avg_per_var_pattern, color = "#fc3d03", linetype = "twodash", size = 1.25)

p1+p2+p6
```

```{r, fig.width=14, fig.height=5}
tree4 <- decision_tree(df, decisions4)
plot(tree4$tree)

result4 <- merge(tree4$df[, c("id", "group_id", "group_size")], df, by = "id")
knitr::kable(leaves_stats(result4, tree4$cols), escape = FALSE)%>%kable_styling()
```
</center>

### *Tree 5*

Splitting variables:

- percentage of missings [5]
- percentage of instances with missings [15]
- percentage of variables with missings [15, 60]

<center>
```{r tree5, fig.width=12, fig.height=3}
decisions5 <- list(
  "per_missings" = 5,
  "per_instances_missings" = c(15),
  "per_variables_missings" = c(15, 60)
)

p1 <- p1+geom_vline(xintercept = decisions5$per_missings, color = "#fc3d03", linetype = "twodash", size = 1.25)
p2 <- p2+geom_vline(xintercept = decisions5$per_instances_missings, color = "#fc3d03", linetype = "twodash", size = 1.25)
p3 <- p3+geom_vline(xintercept = decisions5$per_variables_missings, color = "#fc3d03", linetype = "twodash", size = 1.25)

p1+p2+p3
```

```{r, fig.width=14, fig.height=5}
tree5 <- decision_tree(df, decisions5)
plot(tree5$tree)

result5 <- merge(tree5$df[, c("id", "group_id", "group_size")], df, by = "id")

knitr::kable(leaves_stats(result5, tree5$cols), escape = FALSE)%>%kable_styling()
```
</center>
